The Revisited Hidden Weight Bit Function

### 2025-08-26 paper.tex
Springer version

\def\anon{0} %Set to 1 for anonymous submissions, hides acknowledgements and author names
\def\full{1} %Set to 0 for springer proceedings

\ifnum\full=1
	\documentclass{llncs}
\else
	\documentclass[10pt, runningheads]{llncs}
	\usepackage{times}
\fi

\input{header.tex}

\begin{document}

\setcounter{page}{1}

\title{The Revisited Hidden Weight Bit Function}
\titlerunning{The Revisited Hidden Weight Bit Function}

\ifnum\anon=0
	\author{%
		Pierrick M\'eaux\inst{1},
		Tim Seuré\inst{1}, 
		Deng Tang\inst{2}
	}
	\authorrunning{P.\ M\'eaux, T.\ Seur\'e, D.\ Tang}
	\institute{%
		University of Luxembourg, Luxembourg\\
		\email{pierrick.meaux@uni.lu, tim.seure@uni.lu} 
		\and
		Shanghai Jiao Tong University, Shanghai, China\\
		\email{dengtang@sjtu.edu.cn}
	}
\else
	\author{}
	\institute{}
\fi

\maketitle

\pagestyle{plain}

@@

The Hidden Weight Bit Function (\hwbf{}) has attracted significant attention since its introduction by Bryant in 1991~\cite{IEEE:Bryant91}. It has been regarded as the simplest example of a function whose binary decision diagram has exponential size~\cite{IEEE:Bryant91,TIA:BoLSW99}. The ease of implementing this function across various computational models—owing to its reliance on computing the Hamming weight of the input and applying a simple linear function—combined with its relatively strong cryptographic properties as demonstrated in~\cite{DAM:WCST14} (\eg, balancedness, nonlinearity, degree, and algebraic immunity), has made it a noteworthy candidate for use as a filter function in stream cipher constructions.

Recent developments in stream ciphers have reignited interest in this function. The emergence of new applications for stream ciphers with filter functions on a larger number of variables—such as Hybrid Homomorphic Encryption (\textsf{HHE})~\cite{CCS:NaeLauVai11}—has further emphasized the relevance of such functions. \textsf{HHE} requires Boolean functions that can be efficiently evaluated in an input-oblivious algorithm. For instance, several new binary stream ciphers have been proposed since 2016, including Kreyvium~\cite{JOC:CCFLNP18}, \textsf{FLIP}~\cite{EC:MJSC16}, Rasta~\cite{C:Rasta}, \textsf{FiLIP}~\cite{INDO:MCJS19}, Dasta~\cite{TOSC:BeiLea20} and Fasta~\cite{CTRSA:CidIndRad22}. Currently, designs employing Boolean functions with more than one hundred variables as filters lead to \textsf{HHE} schemes with the best latency~\cite{INDO:HofMeaRic20,CCS:CDPP22,CIC:MeaParPei24,eprint:AGHM24}.

@@

\subsection{Boolean functions and cryptographic parameters}

@@

\begin{definition}[Algebraic Normal Form, degree]\label{def:anf}
	We call Algebraic Normal Form of a Boolean function $f\in\BN$ its unique representation as an element of the ring $\F_2[x_1,\dots,x_n]/(x_1^2+x_1,\dots,x_n^2+x_n)$, and we express it as $f(x_1,\dots,x_n)= \sum_{I \subseteq [1,n]} a_I \left( \prod_{i \in I} x_i \right)$, where $a_I\in \F_2$. The (algebraic) degree of $f$ is defined by $\degg(f)=0$ in case $f$ is the null function, and $\degg(f)=\max\{ |I|:I\subseteq[1,n],a_I=1\}$ otherwise.
\end{definition}

\begin{definition}[Walsh transform]\label{def:walsh_transform}
	The Walsh transform at $a \in \F_2^n$ of a Boolean function $f\in\BN$ restricted to a subset $S\subseteq \F_2^n$ is defined as $\wt{f,S} (a) = \sum_{x\in S} (-1)^{f(x)+a \cdot x}$. The (unrestricted) Walsh transform of $f$ is then defined as $\wt{f}=\wt{f,\F_2^n}$. For any integer $k$, we also set $\wt{f,k}=\wt{f,\Ekn{k}{n}}$.
\end{definition}

@@

\begin{definition}[Fast algebraic immunity, \eg, \cite{EC:ACGKMR06,IEEE:LiuLinPei11,DCC:CarTan15}]\label{def:FAI}
	The fast algebraic immunity of a Boolean function $f\in \Bfn{n}$ is defined to be $\FAI(f)=\min \{ 2 \AI(f), \min\{\degg(g)+ \degg(fg):g\in\BN,1\leq\degg(g)<\AI(f)\}\}$.
\end{definition}

\subsection{Symmetric Functions, \hwbf{} and weightwise degree-$d$ functions}

Recall that the Boolean symmetric functions in $n$ variables are those that are constant on the slice $\Ekn{k}{n}$ for every $k\in [0,n]$.  This class of functions has been thoroughly studied in the context of cryptography, see \eg, \cite{IEEE:Carlet04,IEEE:CanVid05,INDO:BraPre05,DM:SarMai07,IEEE:QFLW09,IEEE:CheLu11,Latin:Meaux19,CCDS:Meaux21,IEEE:CarMea21}. In this article, the symmetric functions that we consider will mainly be the slice indicator function and the majority function.

\begin{definition}[Slice indicator functions]\label{def:slice}
	The indicator function of the slice of weight $k\in[0,n]$ is the function $\phikn kn\in\BN$ defined by $\phikn kn(x)=1$ if and only if $\w(x)=k$.
\end{definition}

@@

\begin{Prop}\label{prop:wwd1}\mbox{}
	\begin{itemize}
		\item (\textit{E.g.}, Theorem $3$ and Theorem $2$ in~\cite{DCC:DalMaiSar06}) The majority function $\mathsf{Maj}_n\in\BN$ satisfies $\NL(\mathsf{Maj}_n)=2^{n-1}-\binom{n-1}{\frac n2}$ and $\degg(\mathsf{Maj}_n)=2^{\lfloor\log_2(n)\rfloor}$ for even $n\geq 2$.
		\item (Theorem $3$ and Theorem $1$ in~\cite{DAM:WCST14}) The \hwbf{} $h\in\BN$ satisfies $\NL(h)=2^{n-1}-2\binom{n-2}{\frac{n-2}{2}}$ for even $n\geq 2$ and $\degg(h)=n-1$ for even $n\geq 4$.
	\end{itemize}
\end{Prop}

\subsection{Krawtchouk polynomials}

@@

\begin{definition}[Krawtchouk polynomials]\label{def:Kraw}
	The $n$-th Krawtchouk polynomial of degree $k\in\Z$ is given by $\kraw{k}{x}{n}=\sum_{j=0}^{k} (-1)^j \binom{x}{j} \binom{n-x}{k-j}$. The Krawtchouk polynomials can alternatively be characterized as the coefficients in the generating function for $(1+z)^{n-x} (1-z)^x$ in the variable $z$, that is, $(1+z)^{n-x} (1-z)^x=\sum_{k\in\Z} \kraw{k}{x}{n} z^k$.
\end{definition}

\section{Revisited \hwbf{} and balancedness}\label{sec:defAndBal}

@@

\subsection{Definition and basic properties}

@@

\begin{proposition}\label{prop:restrWT}
	Let $n\geq 0$ be an even integer, and let $f\in\BN$ be the revisited \hwbf{}. Consider any binary vector $a\in\F_2^n$ and any integer $k\in[1,n]$. Define $c=a+e_k+\sum_{i=n/2+1}^n e_i$. Then the restricted Walsh transform of $f$ satisfies:
	\[
		\wtk{f}{k}(a)=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+n/2}+ c \cdot x}.
	\]
\end{proposition}

\begin{proof}
	We compute:
	\begin{align*}
		\wtk{f}{k}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+a \cdot x}\\
		&=\sum_{x \in \Ekn{k}{n}} (-1)^{x_k + \sum_{i=1}^{n/2} (x_i+1) x_{i+n/2}+ a \cdot x}\\
		&=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+n/2}+\left(x_k+\sum_{i=n/2+1}^{n}x_i+a\cdot x\right)}\\
		&=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+n/2}+ c \cdot x}.\qedhere
	\end{align*}
\end{proof}

@@

\begin{definition}
	Let $n\geq 0$ be an even integer. Then we define the Boolean function $d_n\in\BN$ by:
	\[
		d_n(x)=\sum_{i=1}^{n/2}x_{2i-1}x_{2i}.
	\]
	Further, for an integer $k$ and a binary vector $a\in\F_2^n$, we define:
	\[
		\Dnka{n}{k}{a}=\wtk{d_n}{k}(a)=\sum_{x\in \Ekn{k}{n}} (-1)^{d_n(x) +a\cdot x}.
	\]
\end{definition}

The $\Dnka{n}{k}{a}$ satisfy a recursive relation.

\begin{proposition}\label{prop:recursiveDnka}
Let $n\geq 0$ be an even integer. Then for all integers $k$ and all binary vectors $a\in\F_2^n$, the following hold:
	\begin{itemize}
		\item We have $\Dnka 00a=1$ and $\Dnka 0ka=0$ if $k\neq 0$.
		\item If $n\geq 2$, then for $b=(a_1,\dots,a_{n-2})\in\F_2^{n-2}$, we have:
		\begin{align*}
			\Dnka{n}{k}{a}&=\Dnka{n-2}{k}{b}+\left((-1)^{a_{n-1}} + (-1)^{a_{n}}\right)\Dnka{n-2}{k-1}{b}\\
			&\qquad+ (-1)^{1+a_{n-1}+a_n} \Dnka{n-2}{k-2}{b}.
		\end{align*}
	\end{itemize}
\end{proposition}

\begin{proof}
    We have that $\Dnka 00a=(-1)^0=1$ since $\Ekn 00=\{\varepsilon\}$ for the empty vector $\varepsilon=a\in\F_2^0$. That $\Dnka 0ka=0$ if $k\neq 0$ follows from $\Ekn k0$ being empty if $k\neq 0$. Assume now that $n\geq 2$. By considering binary vectors $x\in\F_2^n$ as $x=(y,x_{n-1},x_n)$ for $y\in\F_2^{n-2}$, we get:
    \begin{align*}
		\Dnka{n}{k}{a}&=\sum_{x\in \Ekn{k}{n}} (-1)^{d_n(x) +a\cdot x}\\
		&=\sum_{y\in \Ekn{k}{n-2}} (-1)^{d_{n-2}(y) +b\cdot y}+
		\sum_{y\in \Ekn{k-1}{n-2}} (-1)^{d_{n-2}(y) +b\cdot y + a_{n-1}}\\
		&\qquad+\sum_{y\in \Ekn{k-1}{n-2}} (-1)^{d_{n-2}(y) +b\cdot y + a_{n}}\\
		&\qquad+\sum_{y\in \Ekn{k-2}{n-2}} (-1)^{d_{n-2}(y) +b\cdot y +1 + a_{n-1}+ a_{n}}\\
		&=\Dnka{n-2}{k}{b}+\left((-1)^{a_{n-1}}+ (-1)^{a_{n}}\right)\Dnka{n-2}{k-1}{b}\\
		&\qquad+ (-1)^{1+a_{n-1}+a_n} \Dnka{n-2}{k-2}{b}.\qedhere
    \end{align*}
\end{proof}

\begin{remark}
    We note that Proposition~\ref{prop:recursiveDnka} gives three different cases depending on the values of the two last elements of $a$:%Then it gives three cases:
	\begin{itemize}
		\item if $a_{n-1}=0=a_n$, then $\Dnka{n}{k}{a}=\Dnka{n-2}{k}{b}+2\Dnka{n-2}{k-1}{b}-\Dnka{n-2}{k-2}{b}$,
		\item if $a_{n-1} \ne a_n$, then $\Dnka{n}{k}{a}=\Dnka{n-2}{k}{b}+\Dnka{n-2}{k-2}{b}$,
		\item if $a_{n-1}=1=a_n$, then $\Dnka{n}{k}{a}=\Dnka{n-2}{k}{b}-2\Dnka{n-2}{k-1}{b}-\Dnka{n-2}{k-2}{b}$.
	\end{itemize}
\end{remark}

\begin{remark}\label{rmk:dependence_dnka}
    The recursive formula of Proposition~\ref{prop:recursiveDnka} gives different cases depending on the values of the pair $(a_{n-1},a_n)$. However, the same reasoning applies to any pair of the form $(a_{2i-1}, a_{2i})$ for $i\in[1,n/2]$. Therefore, the value of $\Dnka{n}{k}{a}$ depends only on the number of pairs $(a_{2i-1}, a_{2i})$ in $a$ being $(0,0)$, $(1,1)$, $(0,1)$ or $(1,0)$. We give the values of $\Dnka{2}{k}{a}$ for $n=2$ in Table~\ref{tab:Dnka}. These values together with Proposition~\ref{prop:recursiveDnka} are sufficient to determine any $\Dnka{n}{k}{a}$.
\end{remark}

\begin{table}
	\scriptsize
	\centering
	\begin{tabular}{|c|c|c|c|}
		\hline
		$a$ & $(0,0)$ & $(0,1)$ & $(1,1)$ \\
		\hline
		$\Dnka{2}{0}{a}$ &$1$&$1$ & $1$\\
		$\Dnka{2}{1}{a}$ &$2$& $0$ & $-2$\\
		$\Dnka{2}{2}{a}$ &$-1$&$1$ & $-1$\\
		\hline
	\end{tabular}
	\mycaption{Values of $\Dnka{2}{k}{a}$.}\label{tab:Dnka}
\end{table}

In the following, we show how the Walsh transform of the revisited \hwbf{} $f\in\BN$ can be written in terms of the $\Dnka{n}{k}{a}$. Then, in Section~\ref{sec:balancedness}, we use Proposition~\ref{prop:recursiveDnka} to determine the balancedness of $f$, and in Section~\ref{sec:genfunc} we study the value of $\Dnka{n}{k}{a}$ using generating functions.

\begin{proposition}\label{prop:WT}
	Let $n\geq 0$ be an even integer, and let $f\in\BN$ be the revisited \hwbf{}. Denote by $\pi:[1,n]\to[1,n]$ the permutation sending the first $n/2$ elements to the odd positions and the $n/2$ last ones to the even positions. Let $a\in\F_2^n$ be a binary vector, and let $b=\pi^{-1}(a)+\sum_{i=1}^{n/2}e_{2i}$. Then the following holds true:
	\[ \wt{f}(a)=1+\sum_{k=1}^n \Dnka{n}{k}{b+\pi(e_k)}.\]
\end{proposition}

\begin{proof}
	For every $k\in[1,n]$, we have by Proposition~\ref{prop:restrWT} that:
	\begin{align*}
		\wtk{f}{k}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+a\cdot x}\\
		&=\sum_{\pi(x) \in \Ekn{k}{n}} (-1)^{f(\pi(x))+a \cdot \pi(x)}\\
		&=\sum_{\pi(x) \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_{\pi(i)} x_{\pi(i+n/2)}+ x_{\pi(k)}+a \cdot \pi(x)+ \sum_{i=1}^{n/2} x_{\pi(i+n/2)}}\\
		&=\sum_{\pi(x) \in \Ekn{k}{n}}(-1)^{d_n(x)+x_{\pi(k)}+\pi^{-1}(a)\cdot x+\sum_{i=1}^{n/2}x_{2i}}\\
		&=\sum_{x\in \Ekn{k}{n}}(-1)^{d_n(x)+(b+\pi(e_k))\cdot x}\\
		&=\Dnka{n}{k}{b+\pi(e_k)}.
	\end{align*}
	We conclude by applying $\wt{f}(a)=1+\sum_{k=1}^n\wtk{f}{k}(a)$.
\end{proof}

@@

In this part, we determine for which values of $n$ the revisited \hwbf{} $f\in\BN$ is balanced, using the expression of its Walsh transform in terms of the $\Dnka{n}{k}{a}$ and the recursive relation of these quantities.

@@

\begin{proposition}\label{proposition:generating_fct}
	For every even integer $n\geq 0$ and every binary vector $a\in\F_2^n$, the following holds:
    \[
        \sum_{k\in\Z}\Dnka nka z^k= P_a(z).
    \]
\end{proposition}

>>>>

\documentclass[runningheads,orivec]{llncs}

\let\proof\relax
\let\endproof\relax

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{microtype}

\newcommand{\AI}{\mathsf{AI}}
\newcommand{\Bfn}[1]{\mathcal{B}_{#1}}
\newcommand{\BN}{\mathcal{B}_n}
\newcommand{\C}{\mathbb{C}}
\newcommand{\dd}{\,\mathrm d}
\newcommand{\degg}{\mathsf{deg}}
\newcommand{\dis}{\mathsf{c_1}}
\newcommand{\Dkna}[3]{\mathsf{D}_{#2,#1}(a)}
\newcommand{\Dtkna}[4]{\mathsf{D}_{#1,#2,#3}(#4)}
\newcommand{\ee}{\mathrm e\mkern1mu}
\newcommand{\eg}{\textit{e.g.}}
\newcommand{\Ekn}[2]{\mathsf{E}_{#1,#2}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\FAI}{\mathsf{FAI}}
\newcommand{\hd}{\mathsf{d_H}}
\newcommand{\hwbf}{\ensuremath{\mathsf{HWBF}}}
\newcommand{\ie}{\textit{i.e.}}
\newcommand{\ii}{\mathrm i\mkern1mu}
\newcommand{\kraw}[3]{\mathsf{K}_{#1}(#2,#3)}
\newcommand{\myurl}[1]{\footnote{\texttt{\url{#1}}}}
\newcommand{\NL}{\mathsf{NL}}
\newcommand{\phikn}[2]{\varphi_{#1,#2}}
\newcommand{\w}{\mathsf{w_H}}
\newcommand{\WD}[2]{\mathcal{WD}^{#1}_{#2}}
\newcommand{\wt}[1]{\mathcal W_{#1}} 
\newcommand{\wtk}[2]{\mathcal{W}_{#1,#2}}
\newcommand{\Z}{\mathbb{Z}}

\let\leq=\leqslant
\let\geq=\geqslant

\title{The Revisited Hidden Weight Bit Function}
\titlerunning{Revisited \hwbf{}}

\author{%
    Pierrick Méaux\inst{1}\orcidID{0000-0001-5733-4341} \and
    Tim Seuré\inst{1}\orcidID{0000-0003-4200-7553} \and %Corresponding author
    Deng Tang\inst{2}\orcidID{0000-0002-8373-9200}
}
\authorrunning{P. Méaux et al.}

\institute{%
    University of Luxembourg, Luxembourg\\
    \email{pierrick.meaux@uni.lu, tim.seure@uni.lu} 
    \and
    Shanghai Jiao Tong University, Shanghai, China\\
    \email{dengtang@sjtu.edu.cn}
}

\begin{document}    
    \maketitle

@@

    The Hidden Weight Bit Function (\hwbf{}) has attracted significant attention since its introduction by Bryant in 1991~\cite{IEEE:Bryant91}. It has been regarded as the simplest example of a function whose binary decision diagram has exponential size~\cite{TIA:BoLSW99,IEEE:Bryant91}. The ease of implementing this function across various computational models—owing to its reliance on computing the Hamming weight of the input and applying a simple linear function—combined with its relatively strong cryptographic properties as demonstrated in~\cite{DAM:WCST14} (\eg, balancedness, nonlinearity, degree, and algebraic immunity), has made it a noteworthy candidate for use as a filter function in stream cipher constructions.

    Recent developments in stream ciphers have reignited interest in this function. The emergence of new applications for stream ciphers with filter functions on a larger number of variables—such as Hybrid Homomorphic Encryption (\textsf{HHE})~\cite{CCS:NaeLauVai11}—has further emphasized the relevance of such functions. \textsf{HHE} requires Boolean functions that can be efficiently evaluated in an input-oblivious algorithm. For instance, several new binary stream ciphers have been proposed since 2016, including Kreyvium~\cite{JOC:CCFLNP18}, \textsf{FLIP}~\cite{EC:MJSC16}, Rasta~\cite{C:Rasta}, \textsf{FiLIP}~\cite{INDO:MCJS19}, Dasta~\cite{TOSC:BeiLea20} and Fasta~\cite{CTRSA:CidIndRad22}. Currently, designs employing Boolean functions with more than one hundred variables as filters lead to \textsf{HHE} schemes with the best latency~\cite{eprint:AGHM24,CCS:CDPP22,INDO:HofMeaRic20,CIC:MeaParPei24}.

@@

    \subsection{Boolean Functions and Cryptographic Parameters}

@@

    \begin{definition}[Algebraic normal form, degree]\label{def:anf}
    	We call algebraic normal form of a Boolean function $f\in\BN$ its unique representation as an element of the ring $\F_2[x_1,\dots,x_n]/(x_1^2+x_1,\dots,x_n^2+x_n)$, and we express it as $f(x_1,\dots,x_n)= \sum_{I \subseteq [1,n]} a_I \left( \prod_{i \in I} x_i \right)$, where $a_I\in \F_2$. The (algebraic) degree of $f$ is defined by $\degg(f)=0$ in case $f$ is the null function, and $\degg(f)=\max\{ |I|:I\subseteq[1,n],a_I=1\}$ otherwise.
    \end{definition}

    \begin{definition}[Walsh transform]\label{def:walsh_transform}
    	The Walsh transform at $a \in \F_2^n$ of $f\in\BN$ restricted to a subset $S\subseteq \F_2^n$ is defined as $\wt{f,S} (a) = \sum_{x\in S} (-1)^{f(x)+a\cdot x}$. The (unrestricted) Walsh transform of $f$ is then defined as $\wt{f}=\wt{f,\F_2^n}$. For any integer $k$, we also set $\wt{f,k}=\wt{f,\Ekn{k}{n}}$.
    \end{definition}

@@

    \begin{definition}[Fast algebraic immunity, \eg, \cite{EC:ACGKMR06,DCC:CarTan15,IEEE:LiuLinPei11}]\label{def:FAI}
    	The fast algebraic immunity of $f\in \Bfn{n}$ is defined as $\FAI(f)=\min \{ 2 \AI(f), \min\{\degg(g)+ \degg(fg):g\in\BN,1\leq\degg(g)<\AI(f)\}\}$.
    \end{definition}

    \subsection{Symmetric Functions, \hwbf{} and Weightwise Degree-$d$ Functions}

    Recall that the Boolean symmetric functions in $n$ variables are those that are constant on the slice $\Ekn{k}{n}$ for every $k\in [0,n]$.  This class of functions has been thoroughly studied in the context of cryptography, see \eg, \cite{INDO:BraPre05,IEEE:CanVid05,IEEE:Carlet04,IEEE:CarMea21,IEEE:CheLu11,Latin:Meaux19,CCDS:Meaux21,IEEE:QFLW09,DM:SarMai07}. In this article, the symmetric functions that we consider will mainly be the slice indicator function and the majority function.

    \begin{definition}[Slice Indicator Functions]\label{def:slice}
    	The indicator function of the slice of weight $k\in[0,n]$ is the function $\phikn kn\in\BN$ defined by $\phikn kn(x)=1$ if and only if $\w(x)=k$.
    \end{definition}

@@

    \begin{property}\label{property:wwd1}\mbox{}
    	\begin{itemize}
    		\item (\textit{E.g.}, Theorem $3$ and Theorem $2$ in~\cite{DCC:DalMaiSar06}) The majority function $\mathsf{Maj}_n\in\BN$ satisfies $\NL(\mathsf{Maj}_n)=2^{n-1}-\binom{n-1}{\frac n2}$ and $\degg(\mathsf{Maj}_n)=2^{\lfloor\log_2(n)\rfloor}$ for even $n\geq 2$.
    		\item (Theorem $3$ and Theorem $1$ in~\cite{DAM:WCST14}) The \hwbf{} $h\in\BN$ satisfies $\NL(h)=2^{n-1}-2\binom{n-2}{\frac{n-2}{2}}$ for even $n\geq 2$ and $\degg(h)=n-1$ for even $n\geq 4$.
    	\end{itemize}
    \end{property}

    \subsection{Krawtchouk Polynomials}

@@

    \begin{definition}[Krawtchouk polynomials]\label{def:Kraw}
    	The $n$-th Krawtchouk polynomial of degree $k\in\Z$ is given by $\kraw{k}{x}{n}=\sum_{j=0}^{k} (-1)^j \binom{x}{j} \binom{n-x}{k-j}$. Alternatively, the Krawtchouk polynomials can be characterized as the coefficients in the generating function for $(1+z)^{n-x} (1-z)^x$ in the variable $z$, that is, $(1+z)^{n-x} (1-z)^x=\sum_{k\in\Z} \kraw{k}{x}{n} z^k$.
    \end{definition}

    \section{Revisited \hwbf{} and Balancedness}\label{sec:defAndBal}

@@

    \subsection{Definition and Basic Properties}

@@

    \begin{proposition}\label{prop:restrWT}
    	Let $n\geq 0$ be an even integer, and let $f\in\BN$ be the revisited \hwbf{}. Consider any binary vector $a\in\F_2^n$ and any integer $k\in[1,n]$. Define $c=a+e_k+\sum_{i=n/2+1}^n e_i$. Then the restricted Walsh transform of $f$ satisfies:
    	\[
    		\wtk{f}{k}(a)=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+n/2}+ c\cdot x}.
    	\]
    \end{proposition}

    \begin{proof}
    	By definition, we have $\wtk{f}{k}(a)=\sum_{x \in \Ekn{k}{n}} (-1)^{x_k + \sum_{i=1}^{n/2} (x_i+1) x_{i+n/2}+ a\cdot x}$. Therefore:
    	\begin{align*}
    		\wtk{f}{k}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+n/2}+\left(x_k+\sum_{i=n/2+1}^{n}x_i+a\cdot x\right)}\\
    		&=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+n/2}+ c\cdot x}.\qedhere
    	\end{align*}
    \end{proof}

@@

    \begin{definition}
    	Let $n\geq 0$ be an even integer. Then we define the Boolean function $d_n\in\BN$ by:
    	\[
    		d_n(x)=\sum_{i=1}^{n/2}x_{2i-1}x_{2i}.
    	\]
    	Further, for an integer $k$ and a binary vector $a\in\F_2^n$, we define:
    	\[
    		\Dkna{n}{k}{a}=\wtk{d_n}{k}(a)=\sum_{x\in \Ekn{k}{n}} (-1)^{d_n(x) +a\cdot x}.
    	\]
    \end{definition}

    The $\Dkna{n}{k}{a}$ satisfy a recursive relation.

    \begin{proposition}\label{prop:recursiveDkna}
    Let $n\geq 0$ be an even integer. Then for all integers $k$ and all binary vectors $a\in\F_2^n$, the following hold:
    	\begin{itemize}
    		\item We have $\Dkna 00a=1$ and $\Dkna 0ka=0$ if $k\neq 0$.
    		\item If $n\geq 2$, then for $b=(a_1,\dots,a_{n-2})\in\F_2^{n-2}$, we have:
    		\begin{align*}
    			\Dkna{n}{k}{a}&=\Dkna{n-2}{k}{b}+\left((-1)^{a_{n-1}} + (-1)^{a_{n}}\right)\Dkna{n-2}{k-1}{b}\\
    			&\qquad+ (-1)^{1+a_{n-1}+a_n} \Dkna{n-2}{k-2}{b}.
    		\end{align*}
    	\end{itemize}
    \end{proposition}

    \begin{proof}
        We have that $\Dkna 00a=(-1)^0=1$ since $\Ekn 00=\{\varepsilon\}$ for the empty vector $\varepsilon=a\in\F_2^0$. That $\Dkna 0ka=0$ if $k\neq 0$ follows from $\Ekn k0$ being empty if $k\neq 0$. Assume now that $n\geq 2$. By considering binary vectors $x\in\F_2^n$ as $x=(y,x_{n-1},x_n)$ for $y\in\F_2^{n-2}$, we get:
        \begin{align*}
    		\Dkna{n}{k}{a}&=\sum_{x\in \Ekn{k}{n}} (-1)^{d_n(x) +a\cdot x}\\
    		&=\sum_{y\in \Ekn{k}{n-2}} (-1)^{d_{n-2}(y) +b\cdot y}+
    		\sum_{y\in \Ekn{k-1}{n-2}} (-1)^{d_{n-2}(y) +b\cdot y + a_{n-1}}\\
    		&\qquad+\sum_{y\in \Ekn{k-1}{n-2}} (-1)^{d_{n-2}(y) +b\cdot y + a_{n}}\\
    		&\qquad+\sum_{y\in \Ekn{k-2}{n-2}} (-1)^{d_{n-2}(y) +b\cdot y +1 + a_{n-1}+ a_{n}}\\
    		&=\Dkna{n-2}{k}{b}+\left((-1)^{a_{n-1}}+ (-1)^{a_{n}}\right)\Dkna{n-2}{k-1}{b}\\
    		&\qquad+ (-1)^{1+a_{n-1}+a_n} \Dkna{n-2}{k-2}{b}.\qedhere
        \end{align*}
    \end{proof}

    \begin{remark}
        We note that Proposition~\ref{prop:recursiveDkna} gives three different cases depending on the values of the two last elements of $a$:
    	\begin{itemize}
    		\item if $a_{n-1}=0=a_n$, then $\Dkna{n}{k}{a}=\Dkna{n-2}{k}{b}+2\Dkna{n-2}{k-1}{b}-\Dkna{n-2}{k-2}{b}$,
    		\item if $a_{n-1} \ne a_n$, then $\Dkna{n}{k}{a}=\Dkna{n-2}{k}{b}+\Dkna{n-2}{k-2}{b}$,
    		\item if $a_{n-1}=1=a_n$, then $\Dkna{n}{k}{a}=\Dkna{n-2}{k}{b}-2\Dkna{n-2}{k-1}{b}-\Dkna{n-2}{k-2}{b}$.
    	\end{itemize}
    \end{remark}

    \begin{remark}\label{rmk:dependence_Dkna}
        The recursive formula of Proposition~\ref{prop:recursiveDkna} gives different cases depending on the values of the pair $(a_{n-1},a_n)$. However, the same reasoning applies to any pair of the form $(a_{2i-1}, a_{2i})$ for $i\in[1,n/2]$. Therefore, the value of $\Dkna{n}{k}{a}$ depends only on the number of pairs $(a_{2i-1}, a_{2i})$ in $a$ being $(0,0)$, $(1,1)$, $(0,1)$ or $(1,0)$. We give the values of $\Dkna{2}{k}{a}$ for $n=2$ in Table~\ref{tab:Dkna}. These values together with Proposition~\ref{prop:recursiveDkna} are sufficient to determine any $\Dkna{n}{k}{a}$.
    \end{remark}

    \begin{table}
    	\scriptsize
    	\centering
        \caption{Values of $\Dkna{2}{k}{a}$.}
        \label{tab:Dkna}
    	\begin{tabular}{|c|c|c|c|}
    		\hline
    		$a$ & $(0,0)$ & $(0,1)$ & $(1,1)$ \\
    		\hline
    		$\Dkna{2}{0}{a}$ &$1$&$1$ & $1$\\
    		$\Dkna{2}{1}{a}$ &$2$& $0$ & $-2$\\
    		$\Dkna{2}{2}{a}$ &$-1$&$1$ & $-1$\\
    		\hline
    	\end{tabular}
    \end{table}

    In the following, we show how the Walsh transform of the revisited \hwbf{} $f\in\BN$ can be written in terms of the $\Dkna{n}{k}{a}$. Then, in Section~\ref{sec:balancedness}, we use Proposition~\ref{prop:recursiveDkna} to determine the balancedness of $f$, and in Section~\ref{sec:genfunc} we study the value of $\Dkna{n}{k}{a}$ using generating functions.

    \begin{proposition}\label{prop:WT}
    	Let $n\geq 0$ be an even integer, and let $f\in\BN$ be the revisited \hwbf{}. Denote by $\pi:[1,n]\to[1,n]$ the permutation sending the first $n/2$ elements to the odd positions and the $n/2$ last ones to the even positions. Let $a\in\F_2^n$ be a binary vector, and let $b=\pi^{-1}(a)+\sum_{i=1}^{n/2}e_{2i}$. Then the following holds true:
    	\[ \wt{f}(a)=1+\sum_{k=1}^n \Dkna{n}{k}{b+\pi(e_k)}.\]
    \end{proposition}

    \begin{proof}
    	For every $k\in[1,n]$, we have by Proposition~\ref{prop:restrWT} that:
    	\begin{align*}
    		\wtk{f}{k}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+a\cdot x}\\
    		&=\sum_{\pi(x) \in \Ekn{k}{n}} (-1)^{f(\pi(x))+a\cdot \pi(x)}\\
    		&=\sum_{\pi(x) \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_{\pi(i)} x_{\pi(i+n/2)}+ x_{\pi(k)}+a\cdot \pi(x)+ \sum_{i=1}^{n/2} x_{\pi(i+n/2)}}\\
    		&=\sum_{\pi(x) \in \Ekn{k}{n}}(-1)^{d_n(x)+x_{\pi(k)}+\pi^{-1}(a)\cdot x+\sum_{i=1}^{n/2}x_{2i}}\\
    		&=\sum_{x\in \Ekn{k}{n}}(-1)^{d_n(x)+(b+\pi(e_k))\cdot x}\\
    		&=\Dkna{n}{k}{b+\pi(e_k)}.
    	\end{align*}
    	We conclude by applying $\wt{f}(a)=1+\sum_{k=1}^n\wtk{f}{k}(a)$.
    \end{proof}

@@

    In this part, we determine for which values of $n$ the revisited \hwbf{} $f\in\BN$ is balanced, using the expression of its Walsh transform in terms of the $\Dkna{n}{k}{a}$ and the recursive relation of these quantities.

@@

    \begin{proposition}\label{proposition:generating_fct}
    	For every even integer $n\geq 0$ and every binary vector $a\in\F_2^n$, the following holds:
        \[
            \sum_{k\in\Z}\Dkna nka z^k= P_a(z).
        \]
    \end{proposition}
